R Notebook for conducting Minimum Detectable Change (MDC) Analysis to design a monitoring strategy for the Upper Yaquina River TMDL Implementation Plan

### 2024-02-15 msc.Rmd
creating function to calculate MDC given numbrer of additional samples

>>>>

## MDC function for Varying N

I used ChatGPT 4.0 to get a starting point for the function

```{r}
calculateMDCWithAdditionalSamples <- function(model, additional_samples, alpha, power) {
  # Extract residuals and X values from the model
  sigma <- sd(model$residuals) # Standard deviation of residuals
  x_values <- model$model[[2]] # Assuming the second column in model data is X
  initial_n <- length(x_values) # Current sample size

  # Calculate the sum of squares of X for the current data
  SS_X_current <- sum((x_values - mean(x_values))^2)

  # Calculate the total number of samples after adding the additional samples
  total_n <- initial_n + additional_samples

  # Assuming the variance of X remains constant, extrapolate SS_X for the total sample size
  # This is a simplification and assumes the distribution of additional X values is similar to the current ones
  SS_X_total <- SS_X_current * (total_n / initial_n)

  # Function to calculate MDC
  calculateMDC <- function(n, alpha, sigma, SS_X, power) {
    df <- n - 2
    t_alpha <- qt(1 - alpha/2, df)
    t_beta <- qt(power, df)
    SE_b <- sigma / sqrt(SS_X)
    MDC <- (t_alpha + t_beta) * SE_b
    return(MDC)
  }

  # Calculate MDC for the total number of samples
  MDC <- calculateMDC(total_n, alpha, sigma, SS_X_total, power)

  return(MDC)
}

No I tried the function for station 12301, which had the largest MDC

mdeol

MDC <- calculateMDCWithAdditionalSamples(model, additional_samples, alpha, power)
print(paste("MDC for", initial_n + additional_samples, "total samples:", MDC))
